\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-08-26 paper/paper.tex
Edits to paper for next Thursday meeting.

\subsubsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.  Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $Y_i$). The likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $Y_1,\ldots,Y_N$ is given by
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

@@

\subsection{Single GLFP Model}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures (1,713), drive-model 14.  The non-parametric Kaplan-Meier estimate plotted below does not appear linear, suggesting different rates of failure in the data. In addition, computer hard drives are known to have a mixture of early and late failures \cite{chan}.  Thus, as a first step, the GLFP model seems like an appropriate parametric model.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{KM14.pdf}
  \caption{Kaplan Meier estimate for drive model 14.  Bands correspond to Greenwood standard errors.}
  \label{fig1}
\end{figure}

@@

The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.  Given the interpretation of $\pi$ as the proportion of defective units, which we would expect to be small, a range primarily covering the left side of 0.50 is reasonable and weakly informative.  The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.  Again, this allows for a wide range of hazards, larger than we would expect to see in the data.  The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$. For $t_{0.2,2}$ the range is $(3.4,16.6)$. Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative. \\

@@

To borrow strength across different populations, we can model the population specific parameters hierarchically, allowing the data to inform the hyperparameters.
%For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:

Let
\begin{equation}
y_{g,i} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes populations. Next we model these parameters themselves as random variables. Let
\begin{equation*}
\sigma_{1g} \ind \op{LN} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
\end{equation*}
\begin{equation*}
\sigma_{2g} \ind \op{LN} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
\end{equation*}
\begin{equation}
\label{eq:hier-model}
\log t_{p_{1}g} \equiv \mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)  \ind \op{N} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
\end{equation}
\begin{equation*}
\log t_{p_{2}g} \equiv \mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)  \ind \op{N} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
\end{equation*}
\begin{equation*}
\op{logit} \pi_g \ind \op{N}(\eta_\pi, \tau_\pi).
\end{equation*}

>>>>

\subsubsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.  Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(t_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(t_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

@@

\subsection{Single Drive-Model Analysis}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, drive-model 14.  The Kaplan-Meier estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures till about 10,000 hours, increases from about 10,000 to 20,000 hours, and then starts to gradually flatten out.  In addition, computer hard drives are known to have a mixture of early and late failures \cite{chan}.  Therefore, there is both an empirical and theoretical justification to apply the GLFP model.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{mod14km.pdf}
  \caption{Kaplan Meier estimate for drive model 14.  Bands correspond to Greenwood standard errors.}
  \label{fig1}
\end{figure}

@@

These priors put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.  Thus, we consider the priors to be relatively uninformative.

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g}, \pi_{g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{f(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-F(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}

Next, we model these parameters themselves as random variables, varying across the different sub-populations.  Let
\begin{equation*}
\sigma_{1g} \ind \op{Lognormal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
\end{equation*}
\begin{equation*}
\sigma_{2g} \ind \op{Lognormal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
\end{equation*}
\begin{equation}
\label{eq:hier-model}
t_{p_{1}g} \equiv \mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)  \ind \op{Lognormal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
\end{equation}
\begin{equation*}
t_{p_{2}g} \equiv \mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)  \ind \op{Lognormal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
\end{equation*}
\begin{equation*}
\op{logit} \pi_g \ind \op{N}(\eta_\pi, \tau_\pi).
\end{equation*}
